### 2014-11-12 doc/thesis/introduction.tex
Some english revisions to the introduction

Dynamically typed languages such as Lua avoid static types in favor of
simplicity and flexibility, because the absence of static types means
that programmers do not need to bother about abstracting types that
should be validated by a type checker.
Instead, dynamically typed languages use run-time \emph{type tags}
to classify the values they compute, so their implementation can use
these tags to perform run-time (or dynamic) type checking
\citep{pierce2002tpl}.

This simplicity and flexibility allow programmers to write code that
might require a quite complex type system to statically type check,
though it may also hide bugs that will be caught only after deployment
if programmers do not properly test their code.
In contrast, static type checking helps programmers detect many
bugs during the development phase.
Static types also provide a conceptual framework that helps
programmers define modules and interfaces that can be combined to
structure the development of programs.

Thus, early error detection and better program structure are two
advantages of static type checking that can lead programmers to
migrate their code from a dynamically typed to a statically
typed language, when their simple scripts evolve into complex programs
\citep{tobin-hochstadt2006ims}.
Dynamically typed languages certainly help programmers during the
beginning of a project, because their simplicity and flexibility
allow quick development and make it easier to change code according to
changing requirements.
However, programmers tend to migrate from dynamically typed to
statically typed code as soon as the project has consolidated its
requirements, because the robustness of static types helps
programmers link requirements to abstractions.
This migration usually involves different languages that have
different syntax and semantics, which usually require a complete
rewrite of existing programs instead of incremental evolution from
dynamic to static types.

Ideally, programming languages should offer programmers the
option to choose between static and dynamic typing:
\emph{optional type systems} \citep{bracha2004pluggable} and
\emph{gradual typing} \citep{siek2006gradual} are two similar
approaches for blending static and dynamic typing in the same
language.
The aim of both approaches is to offer programmers the option
to use type annotations where static typing is needed and to not use
them where dynamic typing is sufficient.
The difference between these two approaches is the way they treat
run-time semantics.
While optional type systems do not affect the run-time semantics,
gradual typing uses run-time checks to ensure that dynamically typed
code does not violate the invariants of statically typed code.

Programmers and researchers sometimes use the term gradual typing
to mean the incremental evolution of dynamically typed code to
statically typed code.
For this reason, gradual typing may also refer to optional type
systems and other approaches that blend static and dynamic typing to
help programmers incrementally migrate from dynamic to static typing
without having to switch to a different language, though all these
approaches differ in the way they handle static and dynamic typing
together.
We use the term gradual typing for referring to the work of
\citet{siek2006gradual}.

In this work we present the design and evaluation of Typed Lua:
an optional type system for Lua that is rich enough to
preserve some of the Lua idioms that programmers are already used to,
but that also includes new constructs that help programmers
structure Lua programs.

Lua is a small imperative language with first-class functions
(with proper lexical scoping) where the only data structure
mechanism is the \emph{table}, an associative array that can
represent arrays, records, maps, modules, objects, etc.
Tables also have syntactic sugar and metaprogramming support
through operator overloading built into the language.
Unlike other scripting languages, Lua has very limited coercion
among different data types.
We believe that designing an optional type system for Lua may
shed some light on optional type systems for scripting languages
in general, as Lua is a small scripting language that shares
some features with other scripting languages such as JavaScript.

Lua prefers to provide mechanisms instead of fixed policies due
to its primary use that has always been as an embedded language
for configuration and extension of other applications.
This means that even features such as a module system and
object orientation are a matter of convention instead of built
into the language.
The result is a fragmented ecosystem of libraries, and different
ideas among Lua programmers on how they should use the language
features, and how they should structure programs.

The lack of standard policies is a challenge for the design of
an optional type system for Lua.
For this reason, we are not relying just on the semantics of
the language to design our type system.
We also run a mostly automated survey of Lua idioms used in a
large corpus of Lua libraries, and it also helped us on the
design of Typed Lua.

So far, Typed Lua is a Lua extension that allows statically typed
code to coexist and interact with dynamically typed code
through optional type annotations, and it also adds standard
constructions for structuring Lua code.
The Typed Lua compiler warns programmers about type errors,
but always generates Lua code that runs in unmodified Lua implementations.
Programmers can enjoy some of the benefits of static types
even without converting existing Lua modules to Typed Lua:
they can export a statically typed interface to a dynamically typed module,
and statically typed users of the module can use the Typed Lua compiler
to check their use of the module.
Thus, implementing an optional type system for Lua offers Lua
programmers one way to obtain most of the advantages of static typing
without compromising the simplicity and flexibility of dynamic typing.
We have an implementation of the Typed Lua compiler that is
available online\footnote{https://github.com/andremm/typedlua}.

Typed Lua intended use is as an application language, and we view
that policies for organizing a program in modules and writing
object-oriented programs should be part of the language and
enforced by its optional type system.
An application language is a programming language that helps
programmers develop applications from scratch until these
applications evolve to complex systems rather than just scripts.
We shall see that Typed Lua introduces the refinement of
tables to support the common way that Lua programmers use
them to represent both modules and objects.

We also believe that Typed Lua help programmers give a more
formal documentation to already existing Lua code, as static types
are also a useful source of documentation in languages that provide
type annotations, because type annotations are always validated by
the type checker and therefore never get outdated.
Thus, programmers can use Typed Lua to define axioms about the
interfaces and types of dynamically typed modules.
We enforce this point by using Typed Lua to statically type
the interface of the Lua standard libraries and other commonly used
Lua libraries, so our compiler can check their usage by Typed Lua
code.

Typed Lua performs a very limited form of local type inference
\citep{pierce2000lti}, as static typing does not necessarily mean
that programmers need to insert type annotations in the code.
Several statically typed languages such as Haskell provide some
amount of type inference that automatically deduces the types of
expressions.
Still, Typed Lua needs a small amount of type annotations due
to the nature of its optional type system.

@@

>>>>

Dynamically typed languages such as Lua avoid static types in favor of
simplicity and flexibility, because the absence of static types means
that programmers do not need to bother with abstracting types that
should be validated by a type checker.
Instead, dynamically typed languages use run-time \emph{type tags}
to classify the values they compute, so their implementation can use
these tags to perform run-time (or dynamic) type checking
\citep{pierce2002tpl}.

This simplicity and flexibility allows programmers to write code that
might require a complex type system to statically type check,
though it may also hide bugs that will be caught only after deployment
if programmers do not properly test their code.
In contrast, static type checking helps programmers detect many
bugs during the development phase.
Static types also provide a conceptual framework that helps
programmers define modules and interfaces that can be combined to
structure the development of programs.

Thus, early error detection and better program structure are two
advantages of static type checking that can lead programmers to
migrate their code from a dynamically typed to a statically
typed language, when their simple scripts evolve into complex programs
\citep{tobin-hochstadt2006ims}.
Dynamically typed languages certainly help programmers during the
beginning of a project, because their simplicity and flexibility
allows quick development and make it easier to change code according to
changing requirements.
However, programmers tend to migrate from dynamically typed to
statically typed code as soon as the project has consolidated its
requirements, because the robustness of static types helps
programmers link requirements to abstractions.
This migration usually involves different languages that have
different syntaxes and semantics, which usually require a complete
rewrite of existing programs instead of incremental evolution from
dynamic to static types.

Ideally, programming languages should offer programmers the
option to choose between static and dynamic typing:
\emph{optional type systems} \citep{bracha2004pluggable} and
\emph{gradual typing} \citep{siek2006gradual} are two similar
approaches for blending static and dynamic typing in the same
language.
The aim of both approaches is to offer programmers the option
to use type annotations where static typing is needed and to avoid
them where dynamic typing is sufficient.
The difference between these two approaches is the way they treat
run-time semantics.
While optional type systems do not affect run-time semantics,
gradual typing uses run-time checks to ensure that dynamically typed
code does not violate the invariants of statically typed code.

Programmers and researchers sometimes use the term \emph{gradual typing}
to mean the incremental evolution of dynamically typed code into
statically typed code.
For this reason, gradual typing may also refer to optional type
systems and other approaches that blend static and dynamic typing to
help programmers incrementally migrate from dynamic to static typing
without having to switch to a different language, though all these
approaches differ in the way they handle static and dynamic typing
together.
We use the term \emph{gradual typing} to refer to the work of
\citet{siek2006gradual}.

In this work we present the design and evaluation of Typed Lua:
an optional type system for Lua that is rich enough to
preserve some of the Lua idioms that programmers are already familiar with,
but that also includes new constructs that help programmers
structure Lua programs.

Lua is a small imperative language with first-class functions
(with proper lexical scoping) where the only data structure
mechanism is the \emph{table} --
an associative array that can represent arrays, records, maps, modules, objects, etc.
Tables also have syntactic sugar and metaprogramming support
through operator overloading built into the language.
Unlike other scripting languages, Lua has very limited coercion
among different data types.

Lua prefers to provide mechanisms instead of fixed policies due
to its primary use as an embedded language for configuration and
extension of other applications.
This means that even features such as a module system and
object orientation are a matter of convention instead of default
language constructs.
The result is a fragmented ecosystem of libraries, and different
ideas among Lua programmers on how they should use the language
features, or how they should structure programs.

The lack of standard policies is a challenge for the design of
an optional type system for Lua.
For this reason, we are not relying entirely on the semantics of
the language to design our type system.
We also run a mostly automated survey of Lua idioms used in a
large corpus of Lua libraries, which also has helped in the design of Typed Lua.

So far, Typed Lua is a Lua extension that allows statically typed
code to coexist and interact with dynamically typed code
through optional type annotations.
In addition, it adds default constructs that programmers can use
to better structure Lua programs.
The Typed Lua compiler warns programmers about type errors,
but always generates Lua code that runs in unmodified Lua implementations.
Programmers can enjoy some of the benefits of static types
even without converting existing Lua modules to Typed Lua --
they can export a statically typed interface to a dynamically typed module,
and statically typed users of the module can use the Typed Lua compiler
to check their use of the module.
Thus, implementing an optional type system for Lua offers Lua
programmers one way to obtain most of the advantages of static typing
without compromising the simplicity and flexibility of dynamic typing.
We have an implementation of the Typed Lua compiler that is
available online\footnote{https://github.com/andremm/typedlua}.

Typed Lua's intended use is as an application language, and
we believe that policies for organizing a program in modules and writing
object-oriented programs should be part of the language and
checked by its optional type system.
An application language is a programming language that helps
programmers develop applications from scratch until these
applications develop into complex systems rather than just scripts.
We shall see that Typed Lua introduces the refinement of
tables to support the common way that Lua programmers use
them to represent both modules and objects.

We also believe that Typed Lua helps programmers give a more
formal documentation to already existing Lua code, as static types
are also a useful source of documentation in languages that provide
type annotations, because type annotations are always validated by
the type checker and therefore never get outdated.
Thus, programmers can use Typed Lua to define axioms about the
interfaces and types of dynamically typed modules.
We enforce this point by using Typed Lua to statically type
the interface of the Lua standard library and other commonly used
Lua libraries, so our compiler can check Typed Lua code that uses
these libraries.

Typed Lua performs a very limited form of local type inference
\citep{pierce2000lti}, as static typing does not necessarily mean
that programmers need to insert type annotations in the code.
Several statically typed languages such as Haskell provide some
amount of type inference that automatically deduces the types of
expressions.
Still, Typed Lua only requires a small amount of type annotations
due to the nature of its optional type system.

@@

Finally, we believe that designing an optional type system for Lua may
shed some light on optional type systems for scripting languages
in general, as Lua is a small scripting language that shares
some features with other scripting languages such as JavaScript.
